---
title: "ESM 206 - Lab 6"
author: "Conner Smith"
date: "11/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(palmerpenguins)
library(broom)
library(equatiomatic)
```

## Rank Based Test Example 

Mann-Whitney U (non-parametric alt. to an unpaired or independent 2-sample t-test).

H0: Ranks (medians) between the two samples are equal 
HA: Samples drawn from populations where ranks are not equal 

### Create samples 

`set.seed()` to create "pseudo-random but reproducible samples"

```{r}
set.seed(1414)
gp_1 <- sample.int(20, size = 15, replace = TRUE)

set.seed(1424)
gp_2 <- sample.int(30, size = 15, replace = TRUE)
```

Take a look at distrib. using `hist()` function 

```{r}
hist(gp_1)
hist(gp_2)
```


Try a t-test so we can compare results:

```{r}
ttest <- t.test(gp_1, gp_2)

# if these data were paired, add argument `paired =  TRUE`
```

This p-value means that, if these samples were drawn from populations with the same mean (i.e. if the null is true), there is a `r round(ttest$p.value,2)` probability of finding sample means *at least as different as these* by random chance.  

[A more comprehensive explanation fo differences between groups ... actual means, data spread, effect size] ... there is no significant difference (t(df)= t-statistic value, p = ). 

### Mann Whitney U rank based test

Use `wilcox.test()` to perform 2-sample rank-based test 

```{r}
mwu <- wilcox.test(gp_1, gp_2)
```

This p-value means that, if these samples were drawn from populations with the same median, there is a `r round(mwu$p.value,2)` probability of finding sample medians *at least as different as these* by random chance.  

... there is no significant difference in median (Mann-Whitney U; U = `r `mwu$statistic`, p = `r round(mwu$p.value,2)`).

## Simple Linear Regression 

Two *continuous quantitative* variables you are trying to understand the relationship between, and you think a linear relationship makes sense.  

Use the `penguins` data exploring flipper length (x) versus body mass (y)

```{r}
# Use view(penguins) to pull up the data 
# Use exploratory scatterplot 

ggplot(data = penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point() +
  geom_smooth(method = "lm")

# body mass(g) = B1 (flipper length (mm))+ B0

```

### Find the mdoel estimates for coefficients 

Use the `lm()` function for linear model 

```{r}
penguin_lm <- lm(body_mass_g ~ flipper_length_mm, data = penguins)
```

`r penguin_lm$coefficients[2]`

Indexing with [2] to access single value from vector 

### Use `broom::tidy()` to get a tidy output from lm()

```{r}
penguin_lm_tidy <- tidy(penguin_lm)

penguin_lm_tidy   
```

### Use broom::glance() to get output of model fit and statistical significance

```{r}
penguin_lm_glance <- glance(penguin_lm)

penguin_lm_glance
```

### Use equatiomatics::exctract_eq() to return the model in LaTeX 

```{r}
extract_eq(model = penguin_lm, use_coefs = TRUE)
```

### Use `plot()` to automatically produce diagnostic plots 

```{r}
plot(penguin_lm)
```

