---
title: "ESM 206 - Lab 6"
author: "Conner Smith"
date: "11/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(palmerpenguins)
library(broom)
library(equatiomatic)
```

## Rank Based Test Example 

Mann-Whitney U (non-parametric alt. to an unpaired or independent 2-sample t-test).

H0: Ranks (medians) between the two samples are equal 
HA: Samples drawn from populations where ranks are not equal 

### Create samples 

`set.seed()` to create "pseudo-random but reproducible samples"

```{r}
set.seed(1414)
gp_1 <- sample.int(20, size = 15, replace = TRUE)

set.seed(1424)
gp_2 <- sample.int(30, size = 15, replace = TRUE)
```

Take a look at distrib. using `hist()` function 

```{r}
hist(gp_1)
hist(gp_2)
```


Try a t-test so we can compare results:

```{r}
ttest <- t.test(gp_1, gp_2)

# if these data were paired, add argument `paired =  TRUE`
```

This p-value means that, if these samples were drawn from populations with the same mean (i.e. if the null is true), there is a `r round(ttest$p.value,2)` probability of finding sample means *at least as different as these* by random chance.  

[A more comprehensive explanation fo differences between groups ... actual means, data spread, effect size] ... there is no significant difference (t(df)= t-statistic value, p = ). 

### Mann Whitney U rank based test

Use `wilcox.test()` to perform 2-sample rank-based test 

```{r}
mwu <- wilcox.test(gp_1, gp_2)
```

This p-value means that, if these samples were drawn from populations with the same median, there is a `r round(mwu$p.value,2)` probability of finding sample medians *at least as different as these* by random chance.  

... there is no significant difference in median (Mann-Whitney U; U = `r `mwu$statistic`, p = `r round(mwu$p.value,2)`).



```{r}

```



